---
title: "Crane Body Size"
author: "Chauncey Gadek and Oona Takano" 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 5
    code_folding: show
    #df_print: paged
    #df_print: kable
    #toc_float: true
      #collapsed: false
      #smooth_scroll: TRUE
    theme: cosmo #spacelab #yeti #united #cosmo
    highlight: tango
  pdf_document:
    df_print: kable
fontsize: 12pt
geometry: margin=0.25in
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{=html}
<style>
/* HTML FORMATTING */
h1, .h1, h2, .h2, h3, .h3, h4, .h4, h5, .h5 {
  margin-top: 25px; /* space before each header */
  font-weight: bold; /* bold headers */
}
</style>
```

```{R, echo=FALSE}
# I set some GLOBAL R chunk options here.
#   (to hide this message add "echo=FALSE" to the code chunk options)
#rm(list =ls (all = TRUE)) #This removes objects from global environ
knitr::opts_chunk$set(echo=F, comment = NA, message = FALSE, warning = FALSE, width = 100)
knitr::opts_chunk$set(fig.align = "center", fig.height = 4, fig.width = 6)

```

# Load Packages

```{R}
pacman::p_load(
AMR,
devtools,
magrittr,
ggpubr,
reshape,
reshape2,
performance,
plyr,
dplyr,
tidyr,
tibble,
car,
rcompanion,
GGally,
Hmisc,
gridExtra,
stats,
gplots,
ggplot2,
ggExtra,
cowplot,
colorspace,
stats4, # Forces knitr to work when it's being wonky
PMCMR, #Allows Kruskal-Wallis post-hocs
effects,
gridExtra,
lattice,
survival,
fmsb,
faraway,
ape,
#wBoot,
ggridges,
boot,
effsize,
plotrix,
colorspace,
patchwork,
ggdist,
# Mapping 
raster,
sp,
rgdal,
prettymapr,
viridis,
rasterVis,
# Modeling packages 
nlme,
lme4,
AICcmodavg,
MuMIn,
reghelper,
lsmeans,
rsq, # get r-squared values from GLM
r2glmm, # for R^2 values from lmer(, and glmer(,
multcompView, # related to multiple comparisons?
jtools, # interaction plots 
interactions, # interaction plots 
broom,
stargazer, # model output tables
ggeffects, # for estimating model predictions from mixed effects models
MCMCglmm,
bayesplot,
rstan,
Rcpp, # required for brms
brms,
magrittr,
tidybayes,
modelr,
hexbin,
ggExtra,
rgl,
readr,
# Phylo packages 
phytools,
ape, 
extrafont)

# To run each time you load rstan
options(mc.cores = parallel::detectCores()) # for core setup 
rstan_options(auto_write = TRUE) # auto save  bare verion of compiled Stan program to HD

#set cores to run for models
n_core <- parallel::detectCores()-1

#Load in functions
source("~/Desktop/R_color_palettes/Gadek_custom_colors.R")
source("~/Desktop/ggplot_themes/ggplot_themes.R")
source("/functions/cbs_model_runners.R")

theme_set(theme_arial_clean())

#setup folder paths for less coding
figures <- paste(getwd(), "/figures/", sep="")
models <- paste(getwd(), "/models/", sep="")
results <- paste(getwd(), "/models/results/", sep="")
```

# Load in data
```{r}

morpho.massive <- read_csv("data/morpho_massive.csv")

```

We have already processed data in the `Data_processing.R` script and run models in the `Run_models.R` script. This .Rmd file is meant for data exploration, statistical tests, and presentation (figure production).

# Descriptive plots
## Mass boxplot
```{r, echo=F}

p <- morpho.massive %>%
  group_by(taxon_name, sex) %>%
  filter(sex %in% c("M", "F"),
         mass >=1800,
         taxon_name %in% c("Antigone canadensis tabida", "Antigone canadensis canadensis")) %>%
  ggplot(., aes(taxon_name, mass, fill = sex)) +
  #geom_violin(trim = F) +
  #geom_box plot(width = 0.1,  position = position_dodge(width = 0.9)) +
  geom_boxplot(width = 0.7,
               alpha = 0.5,
               outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 21,
    size = 2,
    position = position_jitterdodge(jitter.width = 0.15)
  ) +
  scale_x_discrete(labels = c("lesser", "greater")) +
  scale_fill_sex(palette = "sexes") +
  scale_color_sex(palette = "sexes") +
  theme(axis.text.x = element_text(face = "plain"),
        legend.position = c(0.2, 0.85)) +
  labs(x = "", y = "mass (g)")
p
ggsave(
  p,
  filename = paste(
    getwd(),
    "/figures/Mean_mass_subspecies_sex_boxplot_massive_morpho.pdf",
    sep = ""
  ),
  bg = "transparent",
  height = 4.5,
  width = 5,
  units = "in",
  device = cairo_pdf
)

```

## Mass over time
```{r, echo=F}

p <- morpho.massive %>%
  group_by(taxon_name, sex) %>%
  filter(sex %in% c("M", "F"),
         mass >=1800,
         taxon_name %in% c("Antigone canadensis tabida", "Antigone canadensis canadensis")) %>%
  ggplot(., aes(Year, mass, fill = sex, group=sex)) +
  geom_point(
    color = "black",
    shape = 21,
    size = 2
  ) +
  geom_smooth(method = "lm", color="black") +
  scale_fill_sex(palette = "sexes") +
  scale_color_sex(palette = "sexes") +
  theme(axis.text.x = element_text(face = "plain"),
        legend.position = c(0.1, 0.85)) +
  labs(x = "", y = "mass (g)")+
  facet_wrap(.~taxon_name, scales="free_y")
p
ggsave(
  p,
  filename = paste(
    getwd(),
    "/figures/Mean_mass_year_subspecies_sex_massive_morpho.pdf",
    sep = ""
  ),
  bg = "transparent",
  height = 4.5,
  width = 9,
  units = "in",
  device = cairo_pdf
)

```

# Breakpoint regressions
Once possibility I want to explore is that there was a change in measurement tools or measurer at some time(s) in the dataset and these are leading to linear-looking increase. To get at this I will do breakpoint regressions for each explanotory variable.

```{r, `BreakpointLesserMass`}

breaks <- morpho.massive$Year[which(morpho.massive$Year >= 1990 & morpho.massive$Year <= 2010)] #define range of breaks

mse <- numeric(length(breaks))
for(i in 1:length(breaks)){
 piecewise1 <- lm(mass ~ Year*(Year < breaks[i]) + Year*(Year>=breaks[i]), data=morpho.massive)
 mse[i] <- summary(piecewise1)[6]
}
mse <- as.numeric(mse)


plot(mse ~ breaks)

#Breakpoint will show up as lowest point. Looks like 1993?
```




# BRMS Model results
## lesser mass
```{r, `ConditLesserMass`}

ce1 <- conditional_effects(lesser_mass[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "mass (g)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Lesser_cond_eff_mass-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

## Greater mass
```{r, `ConditGreaterMass`}

ce1 <- conditional_effects(greater_mass[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "mass (g)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Greater_cond_eff_mass-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

## lesser wing
```{r, `ConditLesserWing`}

ce1 <- conditional_effects(lesser_wing[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "wing (mm)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Lesser_cond_eff_wing-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

## Greater wing
```{r, `ConditGreaterWing`}

ce1 <- conditional_effects(greater_wing[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "wing (mm)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Greater_cond_eff_wing-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```


## lesser tarsus
```{r, `ConditLesserTarsus`}

ce1 <- conditional_effects(lesser_tarsus[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "tarsus (mm)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Lesser_cond_eff_tarsus-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

## Greater tarsus
```{r, `ConditGreaterTarsus`}

ce1 <- conditional_effects(greater_tarsus[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "tarsus (mm)", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Greater_cond_eff_tarsus-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```


## lesser WT index
```{r, `ConditLesserWTI`}

ce1 <- conditional_effects(lesser_wt_index[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "Wing-Tarsus Index", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Lesser_cond_eff_wing-tarsus-index-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

## Greater WT Index
```{r, `ConditGreaterWTI`}

ce1 <- conditional_effects(greater_wt_index[[3]][[4]])

p <- ggplot(ce1$Year, aes(Year, estimate__)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), fill = "grey") +
  geom_line(color = "black",
            linetype = "dashed",
            linewidth = 1) +
  labs(y = "Wing-Tarsus Index", x = "")
p
ggsave(
  p + theme(),
  filename = paste0(figures, "Greater_cond_eff_wing-tarsus-index-Year.pdf"),
  height = 4,
  width = 4.6,
  units = "in",
  device = cairo_pdf
)

```

# PCA models

```{r, `PCAmodels`}

select.cols.big<-c(4:7, 9:12)
select.cols.little<-c(9:12)

pca.years <- replicate(22, vector("list", 2), simplify = FALSE)

Years <- morpho.massive%>%
  filter(!Year == 2002)%>%
  dplyr::select(Year)%>%
  distinct()%>%
  pull(Year)


for(i in 1:length(pca.years)){ 
  
pca.cols.to.add <-morpho.massive%>%
  filter(sex %in% c("M", "F"),
         Year == Years[i],
         taxon_name %in% c("Antigone canadensis tabida", "Antigone canadensis canadensis"))%>%
  dplyr::select(all_of(select.cols.big))%>%
  na.omit()%>%
  dplyr::select(1:4)


pca.f <-morpho.massive%>%
  filter(sex %in% c("M", "F"),
         Year == Years[i],
         taxon_name %in% c("Antigone canadensis tabida", "Antigone canadensis canadensis"))%>%
  dplyr::select(all_of(select.cols.little))%>%
  na.omit()%>%
  prcomp(., center=T,
         scale=T)


pca.data<- as.data.frame(pca.f$x)

pca.data$species <- pca.cols.to.add$taxon_name
pca.data$sex <- pca.cols.to.add$sex
pca.data$Year <- pca.cols.to.add$Year

pca.years[[i]][[1]] <- pca.data

pca.years[[i]][[2]] <- pca.f

}


names(pca.years) <- Years

pca.score.years <- pca.years[[1]][[1]]

for(i in 2:length(Years)){
  temp <- pca.years[[i]][[1]]
  pca.score.years <- rbind(pca.score.years, temp)
}

```

```{r, `PC1scatter`, out.width=50%}

p <- pca.score.years%>%
  filter(species =="Antigone canadensis tabida")%>%
ggplot(., aes(x=Year, y=PC1, fill=sex))+
  #geom_point(shape=21, size=2)+
  geom_smooth(color="black", method="lm", linetype="dashed")+
  scale_fill_crane("cranes")
p

ggsave(p +theme(), filename = paste0(figures,"Greater_wing_bill_tarsus_culmen_PC1-Year_big_morpho.pdf"), height = 4.6, width = 4.5, units = "in", device = cairo_pdf)



p1 <- pca.score.years%>%
  filter(species =="Antigone canadensis canadensis")%>%
ggplot(., aes(x=Year, y=PC1, fill=sex))+
  #geom_point(shape=21, size=2)+
  geom_smooth(color="black", method="lm", linetype="dashed")+
  scale_fill_crane("cranes")
p1

ggsave(p1 +theme(), filename = paste0(figures,"Lesser_wing_bill_tarsus_culmen_PC1-Year_big_morpho.pdf"), height = 4.6, width = 4.5, units = "in", device = cairo_pdf)

```


```{r, `PC2scatter`, out.width=50%}

p <- pca.score.years%>%
  filter(species =="Antigone canadensis tabida")%>%
ggplot(., aes(x=Year, y=PC2, fill=sex))+
  #geom_point(shape=21, size=2)+
  geom_smooth(color="black", method="lm", linetype="dashed")+
  scale_fill_crane("cranes")
p

ggsave(p +theme(), filename = paste0(figures,"Greater_wing_bill_tarsus_culmen_PC2-Year_big_morpho.pdf"), height = 4.6, width = 4.5, units = "in", device = cairo_pdf)


p <- pca.score.years%>%
  filter(species =="Antigone canadensis canadensis")%>%
ggplot(., aes(x=Year, y=PC2, fill=sex))+
  #geom_point(shape=21, size=2)+
  geom_smooth(color="black", method="lm", linetype="dashed")+
  scale_fill_crane("cranes")
p

ggsave(p +theme(), filename = paste0(figures,"Lesser_wing_bill_tarsus_culmen_PC2-Year_big_morpho.pdf"), height = 4.6, width = 4.5, units = "in", device = cairo_pdf)

```